---
title: "LTBI Bayesian MPES with PREDICT data"
author: "Nathan Green (Imperial College London)"
date: "14/02/2020"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

We will fit the MPES model detailed elsewhere to a PREDICT study data extract.


### Analysis in R

We used vague normal priors for the LTBI prevalence and TB progression.
A prior sensitivity analysis is given at the end of this document.

The data have been grouped so that they are sufficiently coarse for anonymisation and relevant for the DES model.

```{r warning=FALSE, message=FALSE}
library(readr)
library(R2jags)
library(R2WinBUGS)
library(purrr)
library(dplyr)
library(forcats)
```

```{r warning=FALSE, message=FALSE}
data0 <- read_csv(here::here("raw-data", "aggregated_data.csv"),
                  col_types = list(X1 = col_integer(),
                                   age_grp  = col_factor(levels =
                                                           c("(15,35]",
                                                             "(35,55]",
                                                             "(55,100]")),
                                   sex  = col_factor(levels =
                                                       c("Female",
                                                         "Male")),
                                   ethnicity  = col_factor(levels =
                                                             c("White",
                                                               "Black African or Caribbean",
                                                               "South Asian",
                                                               "Other")),
                                   inc_cob_participant2 = col_factor(levels =
                                                                       c('<40',
                                                                         '41-100',
                                                                         '100-300',
                                                                         '>300')),
                                   yearssinceentry_grp = col_factor(levels =
                                                                      c('(0,5]',
                                                                        '(5,10]',
                                                                        '(10,100]')),
                                   prevbcg = col_factor(levels =
                                                          c("No", "Yes")),
                                   reasonforscreening = col_factor(levels =
                                                                     c("Contact",
                                                                       "Migrant")),
                                   pop = col_double(),
                                   tb = col_double(),
                                   ltbi = col_double()))
```

```{r eval=FALSE}
# inspect the raw relationships in the data
library(scales)

data0 <-
  data0 %>% 
  mutate(ltbi_prev = round(ltbi/pop, 3),
         prog_rate = round(tb/ltbi, 3))

# ltbi vs tb counts by grouping
# add some jitter
plot(data0$ltbi + rnorm(nrow(data0),0,0.1), data0$tb + rnorm(nrow(data0),0,0.1),
     col = alpha(as.numeric(data0$reasonforscreening), 0.2), pch = 19,
     xlab = "LTBI", ylab = "TB", main = "reason for screening")

legend("bottomright", legend = c("Contact", "Migrant"),
       col = 1:2, pch = 19, bty = "n")

hist(data0$ltbi_prev[data0$pop > 20],
     breaks = 30,
     main = "",
     xlab = "LTBI prevalence (sample size over 20 people)")
```

```{r}
head(data0)

# dat <- data0[1:100, ]

## aggregate some groups
dat <- 
  data0 %>%
  mutate(
    inc_cob_participant2 = fct_explicit_na(inc_cob_participant2),
    yearssinceentry_grp  = fct_explicit_na(yearssinceentry_grp),
    prevbcg = fct_explicit_na(prevbcg)) %>% 
  group_by(age_grp,
           ethnicity,
           inc_cob_participant2,
           reasonforscreening,
           prevbcg,
           yearssinceentry_grp) %>% 
  summarise(pop = sum(pop),
            tb  = sum(tb),
            ltbi = sum(ltbi))

save(dat, file = "dat.RData")
```

```{r eval=FALSE}
# stratified model

jags_dat_input <-
  list(
    len_gp = nrow(dat),  # number of groups
    Xm  = dat$pop,       # number of migrants/contacts
    Xp  = dat$ltbi,      # number of positive test results
    Xtb = dat$tb         # number of observed active tb cases
  )

jags_dat_input

params <-
  c("sens", "spec",
    "lambda",
    "p_latent")

BUGS_file_name <- "BUGS_code_Xl_fn.txt"
```

```{r eval=TRUE}
# regression model

# transform to levels to integers
# aggregate by covariates of interest
dat_regn <-
  dat %>%
  mutate(
    rfs = as.numeric(reasonforscreening) - 1,
    inc = as.numeric(inc_cob_participant2),
    eth = as.numeric(ethnicity),
    age = as.numeric(age_grp),
    bcg = as.numeric(prevbcg),
    yse = as.numeric(yearssinceentry_grp)
  ) %>%  # transform to level integer
  group_by(rfs,
           inc,
           eth,
           age,
           bcg,
           yse
  ) %>% 
  summarise(pop = sum(pop),
            tb  = sum(tb),
            ltbi = sum(ltbi))

##TODO:
# remove error row
# why is this causing an error?
# pop = ltbi = tb = 1
# maybe its because its prob = 1?
dat_regn <- dat_regn[-235, ]
```

BUGS inputs.

```{r}
# create list of input data
jags_dat_input <-
  list(
    len_gp = nrow(dat_regn),  # number of groups
    len_rfs = max(dat_regn$rfs) + 1,
    len_inc = max(dat_regn$inc),
    len_eth = max(dat_regn$eth),
    len_age = max(dat_regn$age),
    len_bcg = max(dat_regn$bcg),
    len_yse = max(dat_regn$yse),
    rfs = dat_regn$rfs,
    inc = dat_regn$inc,
    eth = dat_regn$eth,
    age = dat_regn$age,
    bcg = dat_regn$bcg,
    yse = dat_regn$yse,
    Xm  = dat_regn$pop,       # number of migrants/contacts
    Xp  = dat_regn$ltbi,      # number of positive test results
    Xtb = dat_regn$tb         # number of observed active tb cases
  )

jags_dat_input

save(jags_dat_input, file = "jags_dat_input.RData")

params <- 
  c("sens", "spec",                # test performance
    "lambda",                      # progression proportion
    "ppred",
    "pred_Xtb", "pred_X_latent",   # posterior predictive distns
    "prior_Xtb", "prior_X_latent"  # prior posterior distns
    # "p_contact", "p_migrant"     # marginalised
  )

BUGS_file_name <- "BUGS_code_regn.txt"
```

```{r}
# n_iter <- 1e6
# n_burnin <- 1e3
# n_thin <- 1e2

n_iter <- 1000#0
n_burnin <- 10#0
n_thin <- 1
```

The BUGS code is

```{r echo=FALSE, comment=NA}
file_loc <- here::here("scripts", BUGS_file_name)

cat(
  readLines(file_loc),
  sep = '\n')
```

```{r message=FALSE, results=FALSE, cache=TRUE}
out <- jags(jags_dat_input,
            parameters.to.save = params,
            model.file = file_loc,
            n.chains = 2,
            n.iter = n_iter,
            n.burnin = n_burnin,
            n.thin = n_thin,
            DIC = TRUE,
            working.directory = here::here("scripts"),
            progress.bar = "text")

save(out, file = "../../data output/out_regn.RData")
```


### References

